---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(forestmangr)
library(dplyr)
```

[![Travis-CI Build Status](https://travis-ci.org/sollano/forestmangr.svg?branch=master)](https://travis-ci.org/sollano/forestmangr)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/forestmangr)](https://cran.r-project.org/package=forestmangr)
[![Coverage Status](https://codecov.io/gh/sollano/forestmangr/branch/master/graph/badge.svg)](https://codecov.io/gh/sollano/forestmangr?branch=master)
[![](https://cranlogs.r-pkg.org/badges/grand-total/forestmangr)](https://cran.rstudio.com/web/packages/forestmangr/index.html)
[![](https://cranlogs.r-pkg.org/badges/forestmangr)](https://cran.rstudio.com/web/packages/forestmangr/index.html)

# forestmangr

Set of functions for processing forest inventory data with methods such as simple random sampling, stratified random sampling and systematic sampling. There are also functions for yield and growth predictions and model fitting, linear and non linear grouped data fitting, and statistical tests.

## Installation
To install the stable CRAN version, use:
```{r eval = FALSE}
install.packages("forestmangr")
```


Or you can install forestmangr from github, for the latest dev version with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("sollano/forestmangr")
```

## Example

```{r}
library(forestmangr)
library(dplyr)
data("exfm16")
head(exfm16)
```

Now, we can fit a model for S estimatation. With `nls_table`, we can fit a non-linear model, extract it's coefficients, and merge it with
the original data in one line. Here we'll use Chapman & Richards model:

```{r}
age_i <- 64
exfm16_fit <- exfm16 %>%
  nls_table(DH ~ b0 * (1-exp(-b1* age))^b2, mod_start = c( b0=23, b1=0.03, b2 = 1.3), output="merge") %>% 
  mutate(site = DH *( ( (1- exp( -b1/age ))^b2 ) / (( 1 - exp(-b1/age_i))^b2 ))) %>% 
  select(-b0,-b1,-b2)
head(exfm16_fit)
```

Now, to fit Clutter's model, we can use the `fit_clutter` function, indicating the DH, B, V, S and Plot variable names:

```{r}
coefs_clutter <- fit_clutter(exfm16_fit, "age", "DH", "B", "V", "site", "plot")
coefs_clutter
```

Now let's say we wanted to do a Simple Random Sampling Forest Inventory, with 20% as a accepted error. 
First, let's load the package and some data:

```{r}
library(forestmangr)
data("exfm2")
data("exfm3")
data("exfm4")
head(exfm3,10)
```
First we should try a pilot inventory, to see if the number of plots sampled is enough for reaching the desired error:

```{r}
sprs(exfm3, "VWB", "PLOT_AREA", "TOTAL_AREA", error = 20, pop = "fin")
```
We can see that we have 10 plots, but 15 more are needed if we want a minimum of 20% error.
The exfm4 data has new samples, that we now can use to run a definitive inventory:

```{r}
sprs(exfm4, "VWB", "PLOT_AREA", "TOTAL_AREA", error = 20, pop = "fin")
```

The desired error was met.

The exfm2 data has a strata variable. Say we wanted to run a SRS inventory for every stand. We can do this with the .groups argument:
```{r}
head(exfm2,10)
sprs(exfm2, "VWB", "PLOT_AREA", "STRATA_AREA",.groups="STRATA", error = 20, pop = "fin")
```

We can also run a stratified random sampling inventory with this data:
```{r}
strs(exfm2, "VWB", "PLOT_AREA", "STRATA_AREA", "STRATA", error = 20, pop = "fin")
```

## Citation
To cite this package in publications, use:

ABNT:

BRAGA S. R., OLIVEIRA, M. L. R., GORGENS, E. B. forestmangr: Functions for Forest Mensuration and Management. R package version 0.9.0, 2018. Disponível em: <https://CRAN.R-project.org/package=forestmangr>

APA:

Sollano Rabelo Braga, Marcio Leles Romarco de Oliveira and Eric Bastos Gorgens (2018). forestmangr: Functions for Forest Mensuration and Management. R package version 0.9.0. https://CRAN.R-project.org/package=forestmangr

## License

This project is licensed under the MIT License - see the [LICENSE](LICENSE) file for details

## Acknowledgments

* This project was developed on the Forest Management Lab, DEF, UFVJM, Diamantina/Minas Gerais - Brazil.

* This project came to be as a mean to make the life of a forestry engeneer a little easier and pratical. 
We'd like to thank everyone at UFVJM that has in anyway helped this project grow.

* We'd like to thank UFVJM, FAPEMIG, CNPq e CAPES for the support.

