---
title: "Phytosociological analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Phytosociological analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
  
```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(tibble.print_min = 6L, tibble.print_max = 6L)
library(forestmangr)
```

For this example we'll use a database of forestry inventories done in the amazon forest, and make a phytosociological analysis of the area.
```{r}
library(forestmangr)
data(exfm20)
data_ex <- exfm20

data_ex
```

First we'll calculate the diversity indexes of the area, with the `species_diversity` function. It just needs the data and column name for species:
```{r}
species_diversity(data_ex, "scientific.name")
```

We can evaluate similarity between plots by the Jaccard index, using the `similarity_matrix` function:
```{r}
similarity_matrix(data_ex, "scientific.name", "transect", index = "Jaccard")
```

We can also generate a dendrogram for this analysis:
```{r}
similarity_matrix(exfm20, "scientific.name", "transect", index = "Jaccard", dendrogram = TRUE, n_groups = 3)
```

To evaluate the level of aggregation among species in the area, we can use the `species_aggreg` function:
```{r}
species_aggreg(data_ex, "scientific.name", "transect")
```

We can also evaluate the horizontal structure of the forest. To do this, we can use the `forest_structure` function:
```{r}
forest_structure(data_ex, "scientific.name", "dbh", "transect", 10000)
```

It's also possible to calculate the vertical and internal structures:
```{r}
forest_structure(data_ex, "scientific.name", "dbh", "transect", 10000, "canopy.pos", "light") 
```

To check if the forest is regulated, we can use the BDq method, with the `bdq_meyer` function:
```{r}
bdq_meyer(data_ex, "transect", "dbh", 1000,licourt_index = 2)
```

With the `diameter_class` function it's possible to divide the data in diameter classes, and get the number of individuals per species in each class:
```{r}
classified <- diameter_class(data_ex,"dbh", "transect", 10000, 10, 10, "scientific.name") 

head(classified)
```

Another way of visualizing this table is to spread the center of class to columns. We can do this with the `cc_to_column` argument:
```{r}
classified <- diameter_class(data_ex,"dbh", "transect", 10000, 10, 10,
               "scientific.name", cc_to_column=TRUE)
head(classified)
```

